Open-source landscape for 3D CSEM modelling

### 2020-05-13 manuscript/Manuscript.tex
Extensive updates to manuscript
main points:

- introduction update
- suggestion of new comprehensive analysis figure type in results part, added also python script to create them in model-block directory
- discussion content extended and rewritten
- conclusion content extended and rewritten
- overall changes: removed redundant statements and reorganized content to fit in either intro, discussion or conclusions part

- RR -

@@

In the last 15 years, the publications with regards to 3D EM modelling grew tremendously, driven by the ever increasing computing power.
One reason why there are so many reports about this topic results from the variety of solution techniques for the systems of linear equations.
They can be distinguished between direct solvers \citep{GEO.09.Streich,
GEO.15.Grayver, GP.14.Chung, GEO.14.Jaysaval, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}.
The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

@@

We continue addressing the topic of open-source developments by introducing and comparing four recent projects with focus on 3D CSEM modeling.
All presented codes are in the Python ecosystem and use either the finite-difference method on structured grids or the finite-element method on unstructured tetrahedral meshes.
As newer package-management systems such as Conda, Docker, or pip significantly help developers to make their codes easily available and installable, each one can be downloaded and installed with a single command.

Furthermore, we deal with the topic of verification.
New codes are generally verified against analytical solutions.
However, analytical and semi-analytical solutions only exist for very simple models.
There is no possibility to check how good a computed response is, expect for comparing results from different modellers.
If different discretisations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy.
We simulate EM fields for a layered background model with vertical transverse isotropy, containing three blocks, and a realistic, complex marine 3D model.
These two model designs as well as the corresponding results from four different codes provide a benchmark for new codes to be compared to and validated with.

It is worth mentioning that there are many more 3D EM codes developed in other fields than geophysics such as communication systems (antennas, radar, satellites), medical imaging, or civil engineering.
There are various open-source codes in these fields too, e.g., Elmer \imycom[DW]{Add 2-3 proper references} (\href{http://www.csc.fi/elmer}{csc.fi/elmer}, couldn't find a publication, just a poster).
While these codes could potentially be used for the same goals as presented here we restrict our review to codes purpose-built and ready without further adjustments for geophysical applications.

\imycom[RR]{I thought a lot about this but I just think the previous paragraph is not required in our introduction and haven't find a location where it fits considering different logical structures. I think it is also not important for this paper to tell "that there are more codes in EM than only geophysics", that's quite obvious. Maybe delete or shorten to 1 sentences to add Elmer and place this sentence somewhere else? I think, an appropriate list of all non-geophysical open-source EM codes should be very long anyway, we just don't know them or at least I'm not familiar with them.}

In the following section we introduce the codes under consideration.
Afterwards, we present the modeling result of the two model cases.
Finally, we discuss, conclude.\mycom[RR]{structure of the paper can be written if we are completeley clear about this and the rest of the introduction.}

@@

The four codes under consideration are, in alphabetical order, \custem
\citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem
\citep{GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}. All four
codes have their user-facing routines written in Python; all of them make heavy
use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy}
\citep{NM.20.Virtanen}. The four of them are “modern” open-source projects,
meaning that they not only come with an open-source license, but they are also
in an online-hosted version-control system with feedback and tracking
possibilities (raising issues, filing pull requests), have extensive online
documentation including many examples, have continuous integration to some
degree, and also importantly are all easily installable, without any user-side
compilations.

@@

\tabl[btp]{codecomparison}{All codes solve the weak form of Maxwell's equations
  in its differential form (DE) under the quasistatic approximation. Note that
  \emg3d is a solver on its own, while the other codes implement third-package
  solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS}
  \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

@@

In this numerical section we compute the responses for two different models to
verify that these different 3D codes yield the same electromagnetic responses.
The first model is a simple one consisting of a layered, anisotropic
background, in which three blocks are embedded, and the second model is a very
complex, realistic model.

\subsection{Block Model}

\cite{GJI.13.Miensopust} published a review of two workshops dealing with the
verification of magnetotelluric forward and inversion codes, hence a similar
exercise to what we do here for CSEM codes. The block model is a derivation of
the \emph{Dublin Test Model 1}, a model from their first workshop. We took the
same layout of the blocks but adjusted the dimensions and resistivities to a
typical, marine CSEM problem, as shown in Figure~\ref{fig:block-model}. We
additionally added a layered background with vertical transverse isotropy
(VTI).
%
\plot*{block-model}{width=.6\textwidth}{
  Sketch of the block model. The layered model consists of an air layer, a
  water layer, a thin top-layer followed by a thick, anisotropic background
  layer, and at the bottom a resistive basement layer. All three blocks are
  embedded in the thick background layer, which has VTI with
  $\lambda=\sqrt{2}$.
}
%

@@

In a first step we compare the layered background to the semi-analytical
solutions of a 1D code, for which we use \empymod \citep{GEO.17.Werthmuller}.
Figures~\ref{fig:results-layered-0} and \ref{fig:results-layered--3} show the
actual responses in the top rows, and the relative error in the bottom rows
for receiver line $y=0\,$km and $y=-3\,$km, respectively (for the layered
background the responses for the two receiver-lines of $y=\pm3$\,km are
identical, so we only show one here).

Figure~\ref{fig:results-layered-0} shows in the top-row the inline responses
for offsets $r>0.5\,$km. As all codes compute the same we only show the
semi-analytical result for the real and imaginary amplitudes, and then in the
bottom row the relative error of the 3D codes. The results show that the codes
are literally yielding the same result, the relative error is for big parts
below 1\,\%, and for almost all parts below a few percent. Some differences are
due to the chosen discretizations. E.g., \emg3d is less precise close to the
200\,m long source. This could be improved by choosing smaller cell sizes, at
the cost of computation time. At large offsets, on the other hand, \emg3d seems
to do a very good job.
%
\plot*{results-layered-0}{width=.8\textwidth}{
   Comparison of the layered model for the inline receivers ($y=0\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   shows the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

The corresponding broadside responses for $y=-3\,$km is shown in
Figure~\ref{fig:results-layered--3}, and the required CPU and RAM is shown in
Table~\ref{tbl:comp-layered}.
%
\plot*{results-layered--3}{width=.8\textwidth}{
   Comparison of the layered model for the broadside receivers ($y=-3\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   show the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

%
\tabl[btp]{comp-layered}{Comparison of used CPU and RAM and of number of cells
  used to discretize the layered model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %{lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#Cells \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

Three resistive blocks are added in the 3.3\,km-thick, anisotropic background
layer, as shown in Figure~\ref{fig:block-model}. They have resistivities of
$\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines),
$\rho=100\,\ohmm$ (thin plate, South-East), $\rho=500\,\ohmm$ (cube,
North-West). As there are no analytical solutions for this model we show the
normalized difference between different codes, instead of the relative error,
where the normalized difference of two responses $p$ and $q$ is given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200 \frac{|p - q|}{|p| + |q|}\ .
  \label{eq:nrmsd}
\end{equation}
%
\imycom[DW]{NRMSD: Comparing all codes to all would make the
figures too cluttered, and I don't think it would add a lot insights. My idea
is therefore to compare \emg3d/\simpeg (the two FV ones), \custem/\petgem (the
two FE ones), and \emg3d/\custem (a cross-comparison). What do you think?}

The results for the three receiver lines $y=-3,0,3\,$km are shown in
Figures~\ref{fig:results-block--3}, \ref{fig:results-block-0}, and
\ref{fig:results-block-3}, respectively, and the corresponding CPU and RAM in
Table~\ref{tbl:comp-block}. What we can conclude from the NRMSD is that similar
codes have a smaller NRMSD than cross-comparison, hence \emg3d and \simpeg are
very similar, as well as are \custem and \petgem.
\mycom[DW]{Write in detail when all results are available}
%
\plot*{results-block--3}{width=.8\textwidth}{
   Comparison of the block model for the broadside receivers at $y=-3\,$km.
   Top row shows the responses and bottom row the NRMSDs (\%)
   between \emg3d--\custem, \custem--\petgem, and \emg3d--\simpeg.
}
%

@@

%
\tabl[btp]{comp-block}{Comparison of used CPU and RAM and of number of cells
  used to discretize the block model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#Cells \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 & 24 & 0.0 & 0 \\
  \emg3d  & 0.0 &  1 & 0.0 & 0 \\
  \petgem & 0.0 & 48 & 0.0 & 0 \\
  \simpeg & 0.0 &    & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

The difference becomes only significant at offsets larger than 8\,km. However,
this is due to a practical difference in the meshing: This simple block model
is advantageous for FV codes, as it consists of rectangular blocks and
horizontal layers, simple geometric objects. While having a very thin, shallow
layer is straight-forward for FV codes, it requires many cells for a FE code.
The thin layer was therefore limited in its horizontal extent
to\mycom[DW]{Raphael, to what x/y exactly?} ???. This is the reason for the
increased NRMSD at large offsets, and it has nothing to do with the accuracy of
the computation itself.

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in
the north-eastern part of the Campos Basin, offshore Brazil, which was
discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well
log data a realistic resistivity model with vertical transverse isotropy (VTI),
Marlim R3D (MR3D), which they released under an open-source license. Using this
model \cite{GEO.19.Correa} computed CSEM data for six frequencies from
0.125\,Hz to 1.25\,Hz, which they also released under an open-source license.
To compute the data they used a state-of-the-art code from the industry
\citep{GEO.07.Maao}. It is therefore an ideal case to verify our open-source
codes against, as it is a complex, realistic model, and the data were computed
by an industry-proofed code. Additionally, that code is a time-domain code,
whereas the four codes under consideration here compute the results all in the
frequency-domain.

@@

The published data set consists of a regular grid of receivers of 20 in
eastern direction by 25 in northern direction, 500 receivers in total, with
1\,km spacing located on the irregular seafloor. 45 source-towlines were
located on the same grid, 50\,m above the seafloor, with shots every 100\,m. In
\cite{GEO.19.Correa} the actual responses for one receiver with the
corresponding East-West inline and one East-West broadside acquisition line are
shown, which is the data we chose for our comparison. The cross-section of the
resistivity model along the chosen inline acquisition line is shown in
Figure~\ref{fig:model-marlim}, together with the acquisition information.
%
\plot*{model-marlim}{width=\textwidth}{
  This figure obviously needs to be improved massively (at least a colorbar).
  What would you like to show from the actual data here?
  A nice figure would  be Figure 2 in \cite{GEO.19.Correa}; however, note that
  the drawn survey line in that figure are plainly wrong. If someone is
  familiar with GMT and happy to make a make that would be great. Also, the
  cross-section is currently heavily vertically exaggerated.
}
%

@@

%
\tabl[btp]{comp-marlim}{Comparison of used CPU and RAM and of number of cells
  used to discretize the Marlim R3D model. \imycom[DW]{I will fill this table
  out once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %{ls[table-format=4.2]rs[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#Cells \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

We compared four different, open-source 3D CSEM modellers by computing
responses for (a) a layered, anisotropic model without and with three resistive
blocks, and (b) a realistic 3D model. The comparison shows that the codes do
yield the same responses. As such we could verify the accuracy of these four
codes.

>>>>

\imycom[RR]{We need to take care to not mix British English with AE, I'm very used to AE, but Dieter started with (at least what I can identify) mainly BE, so I tried to follow (mainly s instead of z and ou instead of o in some words). I hope there are not too many inconsistencies, as probably Lindsey has to fix all these mistakes in the end:-)}

@@

In the last 15 years, the publications with regards to 3D EM modelling grew tremendously, driven by the ever increasing computing power.
One reason why there are so many reports about this topic results from the variety of solution techniques for the systems of linear equations (SLE).
They can be distinguished between direct solvers \citep{GEO.09.Streich,
GEO.15.Grayver, GP.14.Chung, GEO.14.Jaysaval, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}.
The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GP.06.Mulder, GJI.16.Jaysaval}.

@@

\imycom[RR]{i suggest to use also "emph" style for all other package names, not only ours}

We continue addressing the topic of open-source developments with respect to 3D CSEM modeling.
Even though it comprises still the minority of developments, the landscape in this field of research changed in the last five years.
Before, there were only closed-source codes owned by companies or  consortia and individual projects, often without a proper documentation, that could be requested from the authors.
We introduce and compare four recent open-source projects with focus on 3D CSEM modeling.
All presented codes are in the Python ecosystem and use either the FD method on structured grids or the FE method on unstructured tetrahedral meshes.

Furthermore, we deal with the topic of verification.
\cite{GJI.13.Miensopust} presents a review of two workshops dealing with the verification of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data.
Hence, we first conduct a similar exercise for CSEM codes.
We want our efforts to help to ensure the reliability of simulations, not only carried out with the available open-source codes, but also for individual closed-source project.
Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes.
Beyond them, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers.
If different discretisations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy.

We simulate EM fields for a layered background model with vertical transverse isotropy, containing three blocks, and the complex marine Marlim R3D model.
These two model designs as well as the corresponding results from four different codes provide a benchmark for new codes to be compared to and validated with.
In the following section, we introduce the codes under consideration.
Afterwards, we present the considered benchmark cases in detail and present the modeling results of our four codes in terms of accuracy and computational performance.
Beyond that, we extensively discuss important points that control the performance and suitability of our FD and FE codes, including mainly considerations about the mesh design and the choice of solvers. 
We conclude with a summary of our results and a motivation for the EM community to continue not only extending the landscape of open-source codes and also creating a landscape of open-source benchmark models.

\imycom[RR]{Regarding "It is worth mentioning that there are many more 3D EM codes developed in other fields than geophysics such as communication systems (antennas, radar, satellites), medical imaging, or civil engineering.
	There are various open-source codes in these fields too, e.g., Elmer ([DW]: Add 2-3 proper references) (\href{http://www.csc.fi/elmer}{csc.fi/elmer}, couldn't find a publication, just a poster).
	While these codes could potentially be used for the same goals as presented here we restrict our review to codes purpose-built and ready without further adjustments for geophysical applications."

	I thought a lot about this but I just think the previous paragraph is not required in our introduction and haven't find a location where it fits considering different logical structures. I think it is also not important for this paper to tell "that there are more codes in EM than only geophysics", that's quite obvious. Maybe delete or shorten to 1 sentences to add Elmer and place this sentence somewhere else? I think, an appropriate list of all non-geophysical open-source EM codes should be very long anyway, we just don't know them or at least I'm not familiar with them.}

@@

\imycom[RR]{I think we should introduce the term degrees of freedom (dof), equivalent to the size of the SLE, as general measure of our problem sizes. If it is somewhat x*Nedges, or related to Ncells or Nnodes, who cares? We just need this one number for this paper in my opinion, more details in the jupyter notebooks - fine.}

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett}.
All four codes have their user-facing routines written in Python; all of them make heavy use of \texttt{NumPy} \citep{CSE.11.VanDerWalt} and \texttt{SciPy} \citep{NM.20.Virtanen}.
The four of them are “modern” open-source projects, meaning that they come with both, an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests).
All developments comprise an extensive online documentation with many examples and have continuous integration to some degree.
Newer package-management systems such as Conda, Docker, or pip significantly supported us to make our codes easily available and installable without any user-side compilations.
Each one can be downloaded and installed with a single command.

@@

\imycom[RR]{too much info (redundancy to text) for table caption}

\tabl[btp]{codecomparison}{All codes solve the weak form of Maxwell's equations in its differential form (DE) under the quasistatic approximation. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

The customisable electromagnetic modeling Python toolbox custEM was developed for simulating arbitrary complex 3D CSEM geometries with focus on semi-airborne setups, but it supports also land-based, airborne, coastal  or marine environments.
Multiple electric or magnetic field or potential finite-element approaches were implemented as total or secondary field formulations.
The finite-element kernel, including higher order basis functions and parallelisation, relies on the FEniCS project \citep{logg2012, langtangen2016}.
The resulting systems of linear equations are solved with MUMPS \citep{SIAM.01.Amestoy}, which is a very robust but memory consuming choice. 
Primary field solutions are supplied by the COMET \citep{skibbe2020coupled} package.

The toolbox considers generally anisotropic petrophysical properties.
Even though changes of the conductivity is mainly of interest for CSEM modeling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements.
Recently, induced polarization parameters in frequency-domain and three methods for simulating time-domain responses were added to custEM.
The provided meshing tools based on TetGen \citep{si2015tetgen} and functionalities of pyGIMLi \citep{CAG.17.Rucker} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

The custEM toolbox is under steady development and maintenance.
Future enhancements comprise adding iterative solution techniques, provided by FEniCS, as well as further functionalities for automatically incorporating intersecting bodies such as folds or faults in meshes. 
Furthermore, current work is on using custEM for inverse modeling applications.

@@

We computed the responses for two different models to verify that all different 3D codes yield the same electromagnetic responses.
The first simple Block model consists of a layered, anisotropic
background with three embedded blocks.
The second Marlim R3D model is based on a very complex, realistic marine CSEM setup.

\subsection{Block model}

\imycom[RR]{I think the miensopust argument is rather suited to motivate our work in the introduction, i hope its fine that i moved this argument}

The block model is a derivation of the \emph{Dublin Test Model 1} from the first EM modeling workshop described by \cite{GJI.13.Miensopust}.
We took the same layout of the blocks but adjusted the dimensions and resistivities to a typical marine CSEM problem, as shown in Figure~\ref{fig:block-model}.
We additionally added a layered background with vertical transverse isotropy (VTI).
%
\plot*{block-model}{width=.6\textwidth}{
  Sketch of the block model. The layered model consists of an air layer, a
  water layer, a thin top-layer followed by a thick, anisotropic background
  layer, and at the bottom a resistive basement layer. All three blocks are
  embedded in the thick background layer, which has VTI with
  $\lambda=\sqrt{2}$.
}
%

@@

The thin isotropic layer beneath the water requires many tetrahedra for a good-quality discretisation. Therefore, the FE codes custEM and PETGEM used the same tetrahedral mesh with a horizontal layered-earth extent of 30\,km from the origin, surrounded by a ten times larger halfspace-like boundary mesh with water conductivities assigned to the subsurface.
This approximation significantly reduced the problem size compared to considering the layered-earth geometry for the whole computational domain.
\imycom[RR]{Add short information about mesh design with SIMPEG and emg3d, including something like "comparatively coarse discretization in center to minimize computation times" }

In a first step, we compare the layered background to the semi-analytical solutions of a 1D code, for which we use  \empymod\citep{GEO.17.Werthmuller}.
Figures~\ref{fig:results-layered-0} and \ref{fig:results-layered--3} show the actual responses in the top rows, and the relative error in the bottom rows for the receiver lines $y=0\,$km and $y=\pm3\,$km, which are identical.

Figure~\ref{fig:results-layered-0} illustrates the inline responses in the top-row for offsets $r>0.5\,$km. We only show the
semi-analytical result for the real and imaginary amplitudes.
The bottom row displays the relative error of the 3D codes.
The results show that all codes are yielding accurate result with errors of around 1\,\% in most parts and overall below a few percent, disregarding locations affected by sign reversals.
The differences can be mainly attributed to the chosen discretisations.
The FE codes show the strongest misfits at large offsets, most likely caused by the utilized boundary mesh, whereas \emg3d seems to do a very good job\mycom[RR]{maybe reformulate to something like: whereas emg3d appears to be not affected by boundary effects}.
Close to the 200\,m long source, emg3d is less precise which is attributed to comparatively large cell sizes in this region.
%
\plot*{results-layered-0}{width=.8\textwidth}{
   Comparison of the layered model for the inline receivers ($y=0\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   shows the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

\imycom[RR]{I suggest to show y-component for broadside and the errors of the weak component (if not too bad), otherwise only broadside is required to save a plot, because inline should be better for Ex in general.}
The corresponding broadside responses for $y=\pm3\,$km are shown in Figure~\ref{fig:results-layered--3}.
The CPU and RAM requirements are listed in Table~\ref{tbl:comp-layered}.
%
\plot*{results-layered--3}{width=.8\textwidth}{
   Comparison of the layered model for the broadside receivers ($y=-3\,$km).
   Top row shows the semi-analytical responses from \empymod, and bottom row
   show the relative error (\%) of \custem, \emg3d, \petgem, and \simpeg.
}
%

%
\tabl[btp]{comp-layered}{Comparison of used CPU and RAM and of number of cells
  used to discretize the layered model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %{lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

In a second step, three resistive blocks were added in the 3.3\,km-thick, anisotropic background layer as shown in Figure~\ref{fig:block-model}.
They have resistivities of $\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines), $\rho=100\,\ohmm$ (thin plate, South-East), and $\rho=500\,\ohmm$ (cube, North-West).
As there are no analytical solutions for this model we show the normalized difference between different codes, instead of the relative error, where the normalized difference of two responses $p$ and $q$ is given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200 \frac{|p - q|}{|p| + |q|}\ .
  \label{eq:nrmsd}
\end{equation}
%
\imycom[DW]{NRMSD: Comparing all codes to all would make the figures too cluttered, and I don't think it would add a lot insights. My idea is therefore to compare \emg3d/\simpeg (the two FV ones), \custem/\petgem (the two FE ones), and \emg3d/\custem (a cross-comparison). What do you think?}

\imycom[RR]{Agreed. Another suggestion, I solved this visualization problem with a color coded matrix plot, showing 90th percentiles or median values of all errors along the line as meaningful comprehensive number. Look in my thesis - example 4. Comparison for real and imag could be in the lower and upper triangles, or we show absolute magnitudes for layered earth in lower triangle and for block model in upper triangle. With help of this plot, one could easily say if differences are higher for block model and which codes fit to each other well.}

\imycom[RR]{If using the mentioned matrix plots, we would need only one of the figures 4-6, and could show the other data comprehensively.
I added a python script to generate an example of my plotting suggestion in the model-block directory and a new Figure 7 in the manuscript that illustrates what I mean. Of course, any upgrade suggestion for this initial idea is welcome, aside from making it looking nice with labels  and annotations.}

The results for the three receiver lines $y=-3,0,3\,$km are shown in Figures~\ref{fig:results-block--3}, \ref{fig:results-block-0}, and \ref{fig:results-block-3}, respectively, and the corresponding CPU and RAM in Table~\ref{tbl:comp-block}.
We can conclude from the NRMSD that similar codes have a smaller NRMSD than cross-comparison, hence \emg3d and \simpeg are very similar, as well as are \custem and \petgem.
\mycom[DW]{Write in detail when all results are available}
%
\plot*{results-block--3}{width=.8\textwidth}{
   Comparison of the block model for the broadside receivers at $y=-3\,$km.
   Top row shows the responses and bottom row the NRMSDs (\%)
   between \emg3d--\custem, \custem--\petgem, and \emg3d--\simpeg.
}
%

@@

%
\tabl[btp]{comp-block}{Comparison of used CPU and RAM and of number of cells
  used to discretize the block model. \imycom[DW]{I will fill this table out
  once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %lS[table-format=4.2]rS[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 & 24 & 0.0 & 0 \\
  \emg3d  & 0.0 &  1 & 0.0 & 0 \\
  \petgem & 0.0 & 48 & 0.0 & 0 \\
  \simpeg & 0.0 &    & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

The difference becomes only significant at offsets larger than 8\,km. This is due to a practical difference in the meshing: This simple block model
is advantageous for FV codes, as it consists of rectangular blocks and
horizontal layers, simple geometric objects. While having a very thin, shallow
layer is straight-forward for FV codes, it requires many cells for a FE code.
The thin layer was therefore limited in its horizontal extent
to\mycom[DW]{Raphael, to what x/y exactly?} ???. This is the reason for the
increased NRMSD at large offsets, and it has nothing to do with the accuracy of
the computation itself.\mycom[RR]{see imycom below figure 6}

\imycom[RR]{60 x 60 km (30 km in each direction from center). I already added a paragraph about mesh generation in the beginning, I suggest to remove this description here and jsut refer to like "Similar to the layered earth model, the largest errors occur at offsets $>8$\,km due to the differences in mesh design."
We should also mention that relative errors are related to the magnitudes, as real is so much weaker at larger offsets, it is clear that errors increase significantly. Alternatively, we could show absolute magnitude or phase errors such as in the Marlim model.}

\begin{figure}
	\centering
	\includegraphics[width=5cm]{figures/median-stats-block--3.pdf}\hfill
	\includegraphics[width=5cm]{figures/median-stats-block-0.pdf}\hfill
	\includegraphics[width=5cm]{figures/median-stats-block-3.pdf}
	\caption[Median errors]{Overall misfit relations between all codes for the three receiver lines y=-3, y=0 \& y=3 f.l.t.r. Colors indicate the median values of the 101 calculated NRMSD arguments for each line between all codes. Errors for real parts are shown in the lower triangle, for imaginary parts in the upper triangle.}
	\label{medianmisfits}	
\end{figure}

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in the north-eastern part of the Campos Basin, offshore Brazil, which was discovered in 1985.
\cite{BJG.17.Carvalho} created from seismic data and well log data a realistic resistivity model with vertical transverse isotropy (VTI), Marlim R3D (MR3D), which they released under an open-source license. 
\cite{GEO.19.Correa} computed CSEM data for six frequencies from 0.125\,Hz to 1.25\,Hz, released under an open-source license.
To compute the data they used a state-of-the-art code from the industry \citep{GEO.07.Maao}.
It is therefore an ideal case to verify our open-source codes against, as it is a complex, realistic model and the data were computed by an industry-proofed code.\mycom[RR]{I would delete this sentence, not that important and a bit redundant}
Additionally, that code is a time-domain code, whereas the four codes under consideration here compute the results all in the frequency-domain.

@@

The published data set consists of a regular grid of receivers of 20 in eastern direction by 25 in northern direction, 500 receivers in total, with 1\,km spacing located on the irregular seafloor.
45 source-towlines were located on the same grid, 50\,m above the seafloor, with shots every 100\,m. In
\cite{GEO.19.Correa} the actual responses for one receiver with the corresponding East-West inline and one East-West broadside acquisition line are shown, which is the data we chose for our comparison.
The cross-section of the resistivity model along the chosen inline acquisition line is shown in Figure~\ref{fig:model-marlim}, together with the acquisition information.

\imycom[RR]{I used gmt the last time 7 years ago , in between I tried the gmt4py tools, but they did not work well on my machine. I suggest to use two plots next to each other, a similar figure as in the notebook and a simple x-y coord sketch (identical extent as the bird perspective view in the other plot) with all Tx / Rx locations, highlighting the relevant line and Rx for us.}

%
\plot*{model-marlim}{width=\textwidth}{
  This figure obviously needs to be improved massively (at least a colorbar).
  What would you like to show from the actual data here?
  A nice figure would  be Figure 2 in \cite{GEO.19.Correa}; however, note that
  the drawn survey line in that figure are plainly wrong. If someone is
  familiar with GMT and happy to make that would be great. Also, the
  cross-section is currently heavily vertically exaggerated.
}
%

@@

%
\tabl[btp]{comp-marlim}{Comparison of used CPU and RAM and of number of cells
  used to discretize the Marlim R3D model. \imycom[DW]{I will fill this table
  out once we all submitted the final versions of the results.}}{
  \centering
\begin{tabular}{c c c c c}  %{ls[table-format=4.2]rs[table-format=4.2]rl}
  \toprule
  %
  Code & {CPU} & \#Procs & RAM     & \#dof \\
       & {(s)} &         & {(GiB)} &         \\
  \midrule
  %
  \custem & 0.0 &  0 & 0.0 & 0 \\
  \emg3d  & 0.0 &  0 & 0.0 & 0 \\
  \petgem & 0.0 &  0 & 0.0 & 0 \\
  \simpeg & 0.0 &  0 & 0.0 & 0 \\
  %
  \bottomrule
\end{tabular}}%
%

@@

\imycom[RR]{if you prefer ok, but I don't think this is necessary, if everyone adds the own thoughts when he receives the manuscript. Same for conclusions. We need to take care to not mix discussion and conclusions, some of the statements from here belong rather to the conclusions, I hope it's fine that I reorganized the content. Aside from that, I already added all thoughts of mine.}
